### 2004-07-13 Documentation/part1.tex
extended the \GLB command to handle a leading '@' (this is meta character for makeindex and thus has to be escaped by '"'). Added a label for section 7.2
Minor changes
added section about globes put in the fdl part on page two, made the sections of fdl disappear from toc added multiple index support changed the \GLB and \GLBNS commands such, that they prepend 'glb' in the text and add their argument to the api index. makeindex has not be called four times: 	makeindex norm.idx 	makeindex api.idx 	makeindex aedl.idx 	makeindex constants.idx
Adding entries now works with \index{index_name}{index_

\chapter{A \GLOBES\ tour}
\label{chapter:tour}
\index{GLoBES tour}

@@

\begin{table}[tpb]
\begin{center}
\begin{tabular}{p{3.8cm}p{3.8cm}p{7cm}}
\hline
Function & Purpose & Parameters \ra\ Result \\
\hline
\multicolumn{3}{l}{{\bf Systematics only:}} \\
\GLB{glbChiSys} & $\chi^2$ with systematics only  & ({\tt glb\_params in, int exp, int rule}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projections onto axes:}} \\
\GLB{glbChiTheta} & Projection onto $\theta_{13}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{glbChiDelta} & Projection onto $\deltacp$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{glbChiTheta23} & Projection onto $\theta_{23}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{glbChiDm} & Projection onto $\ldm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{glbChiDms} & Projection onto $\sdm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto plane:}} \\
\GLB{glbChiThetaDelta} & Projection onto $\theta_{13}$-$\deltacp$-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto any hyper-plane:}} \\
\GLB{glbChiNP} & Projection onto any $n$-dimensional hyper-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \newline
Needs \GLB{glbSetProjection} before! \\[0.2cm]
\multicolumn{3}{l}{{\bf Localization of degeneracies:}} \\
\GLB{glbChiAll} & (Local) Minimization over all parameters  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\
\hline
\end{tabular}
\end{center}
\caption{\label{tab:stdfunctions} \index{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value with systematics only or systematics and correlations. The parameters {\tt rule} and {\tt exp}
can either be \GLB{GLB\_ALL} for all initialized experiment or the
experiment number ($0$ to \GLB{glb\_num\_of\_exps}-1) for a specific experiment. The format of \GLB{glb\_params} is discussed in detail in \Chapt~\ref{chapt:gettingstarted}. Note that all functions but {\tt ChiSys}
  are using minimizers which have to be initialized with \GLB{glbSetInputErrors} and \GLB{glbSetStartingValues} first.}
\end{table}

@@

\index{Installation}
In this first chapter of the user's manual, we assume that the \GLOBES\ software is readily installed on your computer system. For the installation,
see \App~\ref{app:installation} and the {\tt INSTALL} file in the
software package. We demonstrate how to load pre-defined experiments 
and introduce the basic concepts of \GLOBES . We do not go
into details of the programming language, which means that standard parts
of the program code common to all of the examples in the following chapters are, in general, omitted.
An example of a minimal \GLOBES\ program in C can be found on page~\pageref{ex:c}. Furthermore, the files of the examples in this part can be found in the {\tt Example} subdirectory of your \GLOBES\ distribution. \index{Examples} Since it depends on your installed
configuration, we refer to the {\tt README} file in the \GLOBES\ main directory, and the comments in the Makefile of the {\tt Example} subdirectory for how to compile the example files. Make sure that
the data files (\AEDL\ and supporting files) in the {\tt data} 
subdirectory can be found by the
examples. In case of doubt, you can simply copy the necessary files
into the working directory where the examples are executed.

@@

\example{Using \GLOBES\ with C}{\label{ex:c}
\index{C-Code}
\index{Program}
Here comes the C-code skeleton, which is
(more or less) common to all of our \GLOBES\ examples:
\begin{quote}
{\tt {\footnotesize
\#include <stdio.h> \\
\#include <stdlib.h> \\
\#include <math.h> \\
\#include <string.h> \\
\\
\#include <globes/globes.h> \hspace*{0.5cm} /* Include GLoBES library */ \\
\\
\#include "myio.h" \hspace*{0.5cm} /* Include "housemade" I/O-routines */ \\
\\
/* If filename given, write to file; if empty, to screen: */ \\
char MYFILE[]="testX.dat"; \\
\\
int main(int argc, char *argv[]) \\
\{  \\
\\
 \hspace*{0.5cm} glbInit(argv[0]); \hspace*{0.5cm} /* Initialize GLoBES library */ \\
\\  
  \hspace*{0.5cm} glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\
  \hspace*{1cm} \&glb\_num\_of\_exps); \hspace*{0.5cm} /* Initialize experiment NuFact.glb */\\
\\  
  \hspace*{0.5cm}  /* Initialize housemade output function */\\
  \hspace*{0.5cm} 
   InitOutput(MYFILE,"Format: ... ... ... $\backslash$n"); \\
\\  
  \hspace*{0.5cm} /* Initialize parameter vector(s) */ \\
  \hspace*{0.5cm} glb\_params true\_values = glbAllocParams(); \\
  \hspace*{0.5cm} /* ... */ \\
\\  
 \hspace*{0.5cm} /* Assign: theta12,theta13,theta23,deltacp,dm2solar,dm2atm */ \\
   \hspace*{0.5cm}  
     glbDefineParams(true\_values,\\
     \hspace*{1.5cm}asin(sqrt(0.8))/2,asin(sqrt(0.001))/2,M\_PI/4,M\_PI/2,7e-5,2e-3); \\

@@

\underline{$\beta$-Beams:} & \multicolumn{3}{l}{In preparation} \\[0.1cm]
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:experiments} \index{Experiment files (table)}
\index{{\tt glb}-files} 
Pre-defined experiment prototypes, their filenames (to be used in \GLB{glbInitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
different implementation of the energy threshold function). Note that
some of these experiments are outdated in terms of integrated luminosities,
baseline, fluxes, efficiencies, or other factors.}
\end{table}

\section{Initialization of \GLOBES}
\index{Initialization}

Before one can use \GLOBES , one has to initialize the \GLOBES\
library \GLB{libglobes}: \index{Initialization: \GLOBES\ library}
\begin{function}
\index{Initialization of {\tt libglobes}}
\GLBNS{glbInit} {\tt void glbInit(char *name)} initializes the library {\tt libglobes} and has
to be called in the beginning of each \GLOBES\ program. It takes the
name {\tt name} of the program as a string to initialize the error handling
functions. In many cases, it is sufficient to use the first
argument from the command line as the program name (such as in example on page~\pageref{ex:c}).
\end{function}

\index{Number of experiments}
In principle, the \GLOBES\ user interface can currently handle up to 32 of different long-baseline experiments simultaneously, where the number
of existing experiment definition files can, of course, be unlimited. This means that their $\Delta \chi^2$-values are added {\em after} the minimization over the systematics parameters, and {\em before} any minimization over the oscillation parameters. Note that each experiment
assumes a specific matter density profile for all of its rules, which means
that it makes sense to simulate different operation modes within one
experiment definition, and physically different baselines in different
definitions. For details of the rate computation and
simulation techniques, we refer at this place to \Part~\ref{part:2}. Though
 the simplest case of simulating one experiment may be most often used, 
 using more than one experiments are useful in many cases. For example, combinations of experiments can be tested for
complementarity and competitiveness by equal means within one program.
In general, many \GLOBES\ functions take the experiment number as
a parameter, which runs from $0$ to \GLB{glb\_num\_of\_exps}-1 in the order of their initialization in the program.\footnote{Note that
the global variable {\tt glb\_num\_of\_exps} must not be modified by the
user.} In addition, using the parameter value \GLB{GLB\_ALL} as
experiment number initiates a combined analysis of all loaded experiments.

\index{Experiment list}
\index{Initialization: Experiments}
For storing the experiments, \GLOBES\ uses the initially empty list of experiments \GLB{glb\_experiment\_list}. To add a pre-defined experiment to this list, one can use the function {\tt glbInitExperiment}:
\begin{function}
\index{Experiment initialization}
\GLBNS{glbInitExperiment}
{\tt int glbInitExperiment(char *inf, glb\_exp *in, int *counter)}
 adds a single experiment with the filename {\tt inf} to the list of currently loaded experiments. The {\tt counter} is a pointer to the 
 variable containing the number of experiments, and the experiment {\tt in}
 points to the beginning of the experiment list. The function returns
 zero if it was successful. 
\end{function}
Normally, a typical  call of {\tt glbInitExperiment} is 
\begin{quote}
{\tt glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0],\\  \hspace*{8cm} \&glb\_num\_of\_exps); }
\end{quote}
In this case, the experiment in the file {\tt NuFact.glb} is added to the internal global list of experiments, and the experiment counter is increased. The
experiment then has the number {\tt glb\_num\_of\_exps}-1. The elements
of the experiment list have the type \GLB{glb\_exp}, which the
user will not need to access directly in most cases. The experiment definition files, which usually end with {\tt .glb}, and any
supporting files, are first of all searched in the current directory, and then in the global path variable \GLB{GLB\_PATH} set in your environment.
A list of pre-defined experiment prototypes, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}. If the program cannot find these files, or some of them are syntactically not correct, it will break at this place.

\index{Clear experiment list}
\index{Delete experiments}
One can also remove all experiments from the evaluation list at running
 time:
\begin{function}
\GLBNS{glbClearExperimentList}
{\tt void glbClearExperimentList()} removes all experiments from the internal list and resets all counters.   
\end{function}
Note that changing the number of experiments requires a new initialization
of all parameters of the types \GLB{glb\_params} and \GLB{glb\_projection}
if the number of experiments changes, since these parameter structures internally carry lists for the matter densities of all experiments. Similarly, once should never call {\tt glbAlloc...} before the
experiment initialization.

\begin{table}[tb]
\begin{center}
\begin{tabular}{llp{7cm}}
\hline
Quantities & Examples & Units \\
\hline
Angles & $\theta_{13}$, $\theta_{12}$, $\theta_{23}$, $\deltacp$ & Radians  \\
Mass squared differences & $\sdm$, $\ldm$ & $\mathrm{eV}^2$ \\
Matter densities & $\rho_i$ & $\mathrm{g}/\mathrm{cm}^3$ \\
Baseline lengths & $L_i$ & $\mathrm{km}$ \\
Energies & $E_\nu$ & $\mathrm{GeV}$ \\  
Fiducial masses & $m_{\mathrm{Det}}$ & $\mathrm{t}$ (reactor exp.) or $\mathrm{kt}$ (accelerator exp.), \newline  depends on experiment definition \\
Time intervals & $t_{\mathrm{run}}$ & $\mathrm{yr}$ \\
Source powers & $P_{\mathrm{Source}}$ & Useful parent particle decays/$\mathrm{yr}$ \newline (Neutrino factory, $\beta$-Beam), \newline $\mathrm{GW}$ thermal power (reactor exps.), \newline or
$\mathrm{MW}$ target power (superbeams); \newline
depends on flux definition
 \\
% Integrated luminosities & $m_{\mathrm{Det}} \, t_{\mathrm{run}}$ & $\mathrm{kt \cdot yr}$ \\
Cross sections/E & $\sigma_{\mathrm{CC}}/E$ & $10^{-38} \, \mathrm{cm^2}/\mathrm{GeV^2}$ \\
\hline
\end{tabular}
\mycaption{\label{tab:units} \index{Units in \GLOBES\ (table)} Quantities used in \GLOBES , examples of these quantities, and their standard units in the application software.}
\end{center}
\end{table}

\section{Units in \GLOBES\ and the integrated luminosity}
\label{sec:luminosity}
\index{Integrated luminosity}
\index{Detector mass} \index{Source power} \index{Running time}

While interacting with the user interface of \GLOBES , parameters are transferred to and from the \GLOBES\ library. In \GLOBES , one set of units 
for each type of quantity is used in order to avoid confusion about the definition of individual parameters. \tabl{units} summarizes the units of the most important quantities. In principle, the event rates are
proportional to the product of source power $\times$ target mass $\times$
 running time, which we call ``integrated luminosity''. Since especially the
 definition of the source power depends on the experiment type, the quantities of the three luminosity components
 are not unique and depend on the experiment definition. Usually,
 one uses detector masses in kilotons for beam experiments,
 and detector masses in tons for reactor experiments. Running times
 are normally given in years, where it is often assumed that the 
 experiment runs $100\%$ of the year. Thus, for shorter running periods,
 the running times need to be renormalized. Source powers are
 usually useful parent particle decays per year (neutrino factories,
 $\beta$-beams), target power in mega watts (superbeams), or thermal
 reactor power in giga watts (reactor experiments).
%
Since the pre-defined experiments in \tabl{experiments} are given for specific luminosities, it is useful to read out and change these
 parameters of the individual experiments:
\begin{function}
\GLBNS{glbSetSourcePower}
{\tt void glbSetSourcePower(double power, int exp, int fluxno)}
 sets the source power of experiment number {\tt exp} and flux number
 {\tt fluxno} to {\tt power}. The definition of the source power depends on the experiment type as described above.
\end{function}
\begin{function}
\GLBNS{glbGetSourcePower}
{\tt double glbGetSourcePower(int exp, int fluxno)}
 returns the source power of experiment number {\tt exp} and flux number
 {\tt fluxno}. 
\end{function}
\begin{function}
\GLBNS{glbSetRunningTime}
{\tt void glbSetRunningTime(double time, int exp, int fluxno)}
 sets the running time of experiment number {\tt exp} and flux number
 {\tt fluxno} to {\tt time} years.
\end{function}
\begin{function}
\GLBNS{glbGetRunningTime}
{\tt double glbGetRunningTime(int exp, int fluxno)}
 returns the running time of experiment number {\tt exp} and flux number
 {\tt fluxno}.
\end{function}
\begin{function}
 \GLBNS{glbSetTargetMass}
{\tt void glbSetTargetMass(double mass,int exp)}
sets the fiducial detector mass of experiment number {\tt exp} to {\tt mass} tons or kilotons (depending on the experiment definition).
\end{function}
\begin{function}
 \GLBNS{glbGetTargetMass}
{\tt double glbGetTargetMass(int exp)}
returns the fiducial detector mass of experiment number {\tt exp}.
\end{function}
Thus, these functions also demonstrate how to use the assigned experiment number and others. These numbers run from $0$ to the number of experiments-1, fluxes-1, \etc, where the individual elements are numbered in the order of their appearance. 
 Note that the source power and running time are quantities defined
together with the neutrino flux, whereas the target mass scales the whole
experiment.

\section{Handling oscillation parameter vectors}
\index{Oscillation parameter vectors}

>>>>

\chapter{A \GLOBES\ tour}
\label{chapter:tour}
\index{norm}{GLoBES tour}

@@

\begin{table}[tpb]
\begin{center}
\begin{tabular}{p{3.8cm}p{3.8cm}p{7cm}}
\hline
Function & Purpose & Parameters \ra\ Result \\
\hline
\multicolumn{3}{l}{{\bf Systematics only:}} \\
\GLB{ChiSys} & $\chi^2$ with systematics only  & ({\tt glb\_params in, int exp, int rule}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projections onto axes:}} \\
\GLB{ChiTheta} & Projection onto $\theta_{13}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDelta} & Projection onto $\deltacp$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiTheta23} & Projection onto $\theta_{23}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDm} & Projection onto $\ldm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDms} & Projection onto $\sdm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto plane:}} \\
\GLB{ChiThetaDelta} & Projection onto $\theta_{13}$-$\deltacp$-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto any hyper-plane:}} \\
\GLB{ChiNP} & Projection onto any $n$-dimensional hyper-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \newline
Needs \GLB{SetProjection} before! \\[0.2cm]
\multicolumn{3}{l}{{\bf Localization of degeneracies:}} \\
\GLB{ChiAll} & (Local) Minimization over all parameters  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\
\hline
\end{tabular}
\end{center}
\caption{\label{tab:stdfunctions} \index{norm}{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value with systematics only or systematics and correlations. The parameters {\tt rule} and {\tt exp}
can either be \GLBC{GLB\_ALL} for all initialized experiment or the
experiment number ($0$ to \GLB{\_num\_of\_exps}-1) for a specific experiment. The format of \GLB{\_params} is discussed in detail in \Chapt~\ref{chapt:gettingstarted}. Note that all functions but {\tt ChiSys}
  are using minimizers which have to be initialized with \GLB{SetInputErrors} and \GLB{SetStartingValues} first.}
\end{table}

@@

\index{norm}{Installation}
In this first chapter of the user's manual, we assume that the \GLOBES\ software is readily installed on your computer system. For the installation,
see \App~\ref{app:installation} and the {\tt INSTALL} file in the
software package. We demonstrate how to load pre-defined experiments 
and introduce the basic concepts of \GLOBES . We do not go
into details of the programming language, which means that standard parts
of the program code common to all of the examples in the following chapters are, in general, omitted.
An example of a minimal \GLOBES\ program in C can be found on page~\pageref{ex:c}. Furthermore, the files of the examples in this part can be found in the {\tt Example} subdirectory of your \GLOBES\ distribution. \index{norm}{Examples} Since it depends on your installed
configuration, we refer to the {\tt README} file in the \GLOBES\ main directory, and the comments in the Makefile of the {\tt Example} subdirectory for how to compile the example files. Make sure that
the data files (\AEDL\ and supporting files) in the {\tt data} 
subdirectory can be found by the
examples. In case of doubt, you can simply copy the necessary files
into the working directory where the examples are executed.

@@

\example{Using \GLOBES\ with C}{\label{ex:c}
\index{norm}{C-Code}
\index{norm}{Program}
Here comes the C-code skeleton, which is
(more or less) common to all of our \GLOBES\ examples:
\begin{quote}
{\tt {\footnotesize
\#include <stdio.h> \\
\#include <stdlib.h> \\
\#include <math.h> \\
\#include <string.h> \\
\\
\#include <globes/globes.h> \hspace*{0.5cm} /* Include GLoBES library */ \\
\\
\#include "myio.h" \hspace*{0.5cm} /* Include "housemade" I/O-routines */ \\
\\
/* If filename given, write to file; if empty, to screen: */ \\
char MYFILE[]="testX.dat"; \\
\\
int main(int argc, char *argv[]) \\
\{  \\
\\
 \hspace*{0.5cm} glbInit(argv[0]); \hspace*{0.5cm} /* Initialize GLoBES library */ \\
\\  
  \hspace*{0.5cm} glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\
  \hspace*{1cm} \&glb\_num\_of\_exps); \hspace*{0.5cm} /* Initialize experiment NuFact.glb */\\
\\  
  \hspace*{0.5cm}  /* Initialize housemade output function */\\
  \hspace*{0.5cm} 
   InitOutput(MYFILE,"Format: ... ... ... $\backslash$n"); \\
\\  
  \hspace*{0.5cm} /* Initialize parameter vector(s) */ \\
  \hspace*{0.5cm} glb\_params true\_values = glbAllocParams(); \\
  \hspace*{0.5cm} /* ... */ \\
\\  
 \hspace*{0.5cm} /* Assign: theta12,theta13,theta23,deltacp,dm2solar,dm2atm */ \\
   \hspace*{0.5cm}  
     glbDefineParams(true\_values,\\
     \hspace*{1.5cm}asin(sqrt(0.8))/2,asin(sqrt(0.001))/2,M\_PI/4,M\_PI/2,7e-5,2e-3); \\

@@

\underline{$\beta$-Beams:} & \multicolumn{3}{l}{In preparation} \\[0.1cm]
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:experiments} \index{norm}{Experiment files (table)}
\index{norm}{{\tt glb}-files} 
Pre-defined experiment prototypes, their filenames (to be used in \GLB{InitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
different implementation of the energy threshold function). Note that
some of these experiments are outdated in terms of integrated luminosities,
baseline, fluxes, efficiencies, or other factors.}
\end{table}

\section{Initialization of \GLOBES}
\index{norm}{Initialization}

Before one can use \GLOBES , one has to initialize the \GLOBES\
library \index{norm}{libglobes@{\tt libglobes}}: \index{norm}{Initialization!\GLOBES\ library}
\begin{function}
\index{norm}{Initialization!libglobes@{\tt libglobes}}
\GLBNS{Init} {\tt void glbInit(char *name)} initializes the library {\tt libglobes} and has
to be called in the beginning of each \GLOBES\ program. It takes the
name {\tt name} of the program as a string to initialize the error handling
functions. In many cases, it is sufficient to use the first
argument from the command line as the program name (such as in example on page~\pageref{ex:c}).
\end{function}

\index{norm}{Experiment!number of}
In principle, the \GLOBES\ user interface can currently handle up to 32 of different long-baseline experiments simultaneously, where the number
of existing experiment definition files can, of course, be unlimited. This means that their $\Delta \chi^2$-values are added {\em after} the minimization over the systematics parameters, and {\em before} any minimization over the oscillation parameters. Note that each experiment
assumes a specific matter density profile for all of its rules, which means
that it makes sense to simulate different operation modes within one
experiment definition, and physically different baselines in different
definitions. For details of the rate computation and
simulation techniques, we refer at this place to \Part~\ref{part:2}. Though
 the simplest case of simulating one experiment may be most often used, 
 using more than one experiments are useful in many cases. For example, combinations of experiments can be tested for
complementarity and competitiveness by equal means within one program.
In general, many \GLOBES\ functions take the experiment number as
a parameter, which runs from $0$ to \GLB{\_num\_of\_exps}-1 in the order of their initialization in the program.\footnote{Note that
the global variable {\tt glb\_num\_of\_exps} must not be modified by the
user.} In addition, using the parameter value \GLBC{GLB\_ALL} as
experiment number initiates a combined analysis of all loaded experiments.

\index{norm}{Experiment!list}
\index{norm}{Initialization!experiments}
For storing the experiments, \GLOBES\ uses the initially empty list of experiments \GLB{\_experiment\_list}. To add a pre-defined experiment to this list, one can use the function {\tt glbInitExperiment}:
\begin{function}
\index{norm}{Experiment initialization}
\GLBNS{InitExperiment}
{\tt int glbInitExperiment(char *inf, glb\_exp *in, int *counter)}
 adds a single experiment with the filename {\tt inf} to the list of currently loaded experiments. The {\tt counter} is a pointer to the 
 variable containing the number of experiments, and the experiment {\tt in}
 points to the beginning of the experiment list. The function returns
 zero if it was successful. 
\end{function}
Normally, a typical  call of {\tt glbInitExperiment} is 
\begin{quote}
{\tt glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0],\\  \hspace*{8cm} \&glb\_num\_of\_exps); }
\end{quote}
In this case, the experiment in the file {\tt NuFact.glb} is added to the internal global list of experiments, and the experiment counter is increased. The
experiment then has the number {\tt glb\_num\_of\_exps}-1. The elements
of the experiment list have the type \GLB{\_exp}, which the
user will not need to access directly in most cases. The experiment definition files, which usually end with {\tt .glb}, and any
supporting files, are first of all searched in the current directory, and then in the global path variable \GLBC{GLB\_PATH} set in your environment.
A list of pre-defined experiment prototypes, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}. If the program cannot find these files, or some of them are syntactically not correct, it will break at this place.

\index{norm}{Clear experiment list}
\index{norm}{Delete experiments}
One can also remove all experiments from the evaluation list at running
 time:
\begin{function}
\GLBNS{ClearExperimentList}
{\tt void glbClearExperimentList()} removes all experiments from the internal list and resets all counters.   
\end{function}
Note that changing the number of experiments requires a new initialization
of all parameters of the types \GLB{\_params} and \GLB{\_projection}
if the number of experiments changes, since these parameter structures internally carry lists for the matter densities of all experiments. Similarly, once should never call {\tt glbAlloc...} before the
experiment initialization.

\begin{table}[tb]
\begin{center}
\begin{tabular}{llp{7cm}}
\hline
Quantities & Examples & Units \\
\hline
Angles & $\theta_{13}$, $\theta_{12}$, $\theta_{23}$, $\deltacp$ & Radians  \\
Mass squared differences & $\sdm$, $\ldm$ & $\mathrm{eV}^2$ \\
Matter densities & $\rho_i$ & $\mathrm{g}/\mathrm{cm}^3$ \\
Baseline lengths & $L_i$ & $\mathrm{km}$ \\
Energies & $E_\nu$ & $\mathrm{GeV}$ \\  
Fiducial masses & $m_{\mathrm{Det}}$ & $\mathrm{t}$ (reactor exp.) or $\mathrm{kt}$ (accelerator exp.), \newline  depends on experiment definition \\
Time intervals & $t_{\mathrm{run}}$ & $\mathrm{yr}$ \\
Source powers & $P_{\mathrm{Source}}$ & Useful parent particle decays/$\mathrm{yr}$ \newline (Neutrino factory, $\beta$-Beam), \newline $\mathrm{GW}$ thermal power (reactor exps.), \newline or
$\mathrm{MW}$ target power (superbeams); \newline
depends on flux definition
 \\
% Integrated luminosities & $m_{\mathrm{Det}} \, t_{\mathrm{run}}$ & $\mathrm{kt \cdot yr}$ \\
Cross sections/E & $\sigma_{\mathrm{CC}}/E$ & $10^{-38} \, \mathrm{cm^2}/\mathrm{GeV^2}$ \\
\hline
\end{tabular}
\mycaption{\label{tab:units} \index{norm}{Units in \GLOBES\ (table)} Quantities used in \GLOBES , examples of these quantities, and their standard units in the application software.}
\end{center}
\end{table}

\section{Units in \GLOBES\ and the integrated luminosity}
\label{sec:luminosity}
\index{norm}{Integrated luminosity}
\index{norm}{Detector mass} \index{norm}{Source power} \index{norm}{Running time}

While interacting with the user interface of \GLOBES , parameters are transferred to and from the \GLOBES\ library. In \GLOBES , one set of units 
for each type of quantity is used in order to avoid confusion about the definition of individual parameters. \tabl{units} summarizes the units of the most important quantities. In principle, the event rates are
proportional to the product of source power $\times$ target mass $\times$
 running time, which we call ``integrated luminosity''. Since especially the
 definition of the source power depends on the experiment type, the quantities of the three luminosity components
 are not unique and depend on the experiment definition. Usually,
 one uses detector masses in kilotons for beam experiments,
 and detector masses in tons for reactor experiments. Running times
 are normally given in years, where it is often assumed that the 
 experiment runs $100\%$ of the year. Thus, for shorter running periods,
 the running times need to be renormalized. Source powers are
 usually useful parent particle decays per year (neutrino factories,
 $\beta$-beams), target power in mega watts (superbeams), or thermal
 reactor power in giga watts (reactor experiments).
%
Since the pre-defined experiments in \tabl{experiments} are given for specific luminosities, it is useful to read out and change these
 parameters of the individual experiments:
\begin{function}
\GLBNS{SetSourcePower}
{\tt void glbSetSourcePower(double power, int exp, int fluxno)}
 sets the source power of experiment number {\tt exp} and flux number
 {\tt fluxno} to {\tt power}. The definition of the source power depends on the experiment type as described above.
\end{function}
\begin{function}
\GLBNS{GetSourcePower}
{\tt double glbGetSourcePower(int exp, int fluxno)}
 returns the source power of experiment number {\tt exp} and flux number
 {\tt fluxno}. 
\end{function}
\begin{function}
\GLBNS{SetRunningTime}
{\tt void glbSetRunningTime(double time, int exp, int fluxno)}
 sets the running time of experiment number {\tt exp} and flux number
 {\tt fluxno} to {\tt time} years.
\end{function}
\begin{function}
\GLBNS{GetRunningTime}
{\tt double glbGetRunningTime(int exp, int fluxno)}
 returns the running time of experiment number {\tt exp} and flux number
 {\tt fluxno}.
\end{function}
\begin{function}
 \GLBNS{SetTargetMass}
{\tt void glbSetTargetMass(double mass,int exp)}
sets the fiducial detector mass of experiment number {\tt exp} to {\tt mass} tons or kilotons (depending on the experiment definition).
\end{function}
\begin{function}
 \GLBNS{GetTargetMass}
{\tt double glbGetTargetMass(int exp)}
returns the fiducial detector mass of experiment number {\tt exp}.
\end{function}
Thus, these functions also demonstrate how to use the assigned experiment number and others. These numbers run from $0$ to the number of experiments-1, fluxes-1, \etc, where the individual elements are numbered in the order of their appearance. 
 Note that the source power and running time are quantities defined
together with the neutrino flux, whereas the target mass scales the whole
experiment.

\section{Handling oscillation parameter vectors}
\index{norm}{Oscillation!parameter vectors}

Before we can set the simulated event rates or access any oscillation
parameters, we need to become familiar with the concept \GLOBES\ uses
for oscillation parameters. In order to transfer sets of oscillation
parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$, the parameter type \GLB{\_params\_type} is used. In general, pointers to this structure are of the type \GLB{\_params}, and are often
transferred to and from \GLOBES\ functions. Therefore, the memory space
for these vectors has to be reserved before they can be used, and it has
to be returned afterwards. \GLOBES\ functions usually use the pointers
of the type {\tt glb\_params} for the input or output to the functions. As
an input parameter, the pointer has to be created and point towards a valid
parameter structure, where the oscillation parameters are read from. As
an output parameter, the pointer has to be created, too, and point towards
a structure which will contain the return values will be written to.
This parameter transfer concept seems to be very sophisticated, but, as we
will see in the next chapters, it hides a lot of complicated parameter 
mappings which otherwise need to be done by the user. For example, not
only the oscillation parameters are stored in the pointer structure, but also information on the matter
densities of all of the initialized experiments. Since \GLOBES\ treats
the matter density as a free parameter known with some external precision
to include matter density uncertainties,
the minimizers also use fit values and external errors for the matter
densities of all loaded experiments. More precisely, the matter density
profile of each experiment $i$ is multiplied by a scaling factor
$\hat{\rho_i}$, \index{norm}{Matter density scaling factor}
\index{norm}{Matter density profile} 
which is stored in the density information of {\tt glb\_params}. 
Each of
these scaling factors has $1.0$ as pre-defined value. Since it
is in most cases not necessary to change this value, the user does
not need to take care of it. For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$. For a matter density profile, it acts as an overall normalization factor: The matter density in each layer is multiplied by this factor. In most cases one wants to take a scaling factor of $1.0$ here, which simply means taking the matter density profile as it is given in the experiment definition. For the treatment of correlations, however,
an external precision of the scaling factor might be used to include the
correlations with the matter density uncertainty. For more details, we
refer to the following sections.
Note that the {\tt glb\_params}
structures must not be initialized before all experiments are loaded,
since the number of matter densities can only be determined after the
experiments are initialized. 
Similarly, any change in the number of experiments requires that the
parameter structures be re-initialized, \ie, destroyed and created again.
